Case for Support

### 2021-11-12 detailsOfResource.tex
Added

\documentclass[a4paper,11point]{article}
\usepackage[margin=2cm]{geometry}

%% Fonts, etc.

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage{microtype}      % fine pdf font control

\usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
\usepackage{xcolor} % -- loaded by tikz
% \usepackage{soul} % \ul \st \hl

%% Maths
% \usepackage{amssymb,amsmath}
% \usepackage{algorithm,algpseudocode}
% \usepackage{mathtools}
% \usepackage{mathalfa}
% \usepackage{bbm}
% \usepackage{mathrsfs}
% \usepackage{gensymb}
% \input m.macros
% \input m.symbols

%% Other typsetting
% \usepackage{booktabs}       % professional-quality tables
% \usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{enumitem}
https://www.overleaf.com/project/61853ad43cf7a80aa77cfc1d
%% Graphicshttps://www.overleaf.com/project/61853ad43cf7a80aa77cfc1d
% \usepackage{graphicx,tikz}
% \usetikzlibrary{positioning}
% \usetikzlibrary{fit}
% \usetikzlibrary{calc}

%% Floats
% \usepackage{wrapfig}

%% Margin bubbles
% \usepackage{mbubble}
% \mbubblestyle{ms}{font=\tiny,label=ms,color=blue}
% \newcommand{\comment}[1]{}

%% Sections
% \setcounter{secnumdepth}{-2}
% \usepackage[small]{titlesec} % format section titles
% \titleformat{command}[shape]{format}{label}{sep}{before-code}[after-code]

% \usepackage{caption,subcaption}
% \DeclareCaptionSubType[Alph]{figure}
% \captionsetup[subfigure]{labelformat=simple,labelfont={sf,bf,large},textfont={sf}}
% \renewcommand{\thesubfigure}{{\sffamily\alph{subfigure}}}

%% Spacing
% \usepackage[doublespacing]{setspace}
% \parindent 0pt
% \parskip 1ex

%% Citations, refs, hyperlinks
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{natbib}
\usepackage{apalike}
\usepackage{graphicx}
% \usepackage{natbib}
% \usepackage[capitalize]{cleveref}
% \crefname{equation}{}{}
% \crefname{Equation}{Equation}{Equations}

%% Title

\title{}
\author{}
% \usepackage{filemod} % access to file modification times
% \date{\Filemodtoday{\jobname}}  % replace \jobname if compiling from wrapper

% \usepackage{titling} % modify title(page) appearance
%\pretitle{\begin{center}\LARGE}
%\posttitle{\par\end{center}\vskip 0.5em}
%\preauthor{\begin{center} \large \lineskip 0.5em \begin{tabular}[t]{c}}
%\postauthor{\end{tabular}\par\end{center}}
%\predate{\begin{center}\large}
%\postdate{\par\end{center}}

\def\ii#1\par{{\color{blue!40}\sl #1}\par}
\def\iibf#1\par{{\color{blue!40}\sl\bfseries #1}\par}
\def\iitem#1\par{\ii\begin{itemize}[nosep]\item #1 \end{itemize}\par}

\begin{document}

\noindent\textbf{Notes}

(This page is not for submission)

Required items:
\begin{itemize}  
  \item case for support (up to eight pages)
  \item justification of resources (two pages)
  \item data management plan (DMP) (up to three pages, data management plan template)
        if you opt not to use the template for your plan, the topics listed in the template must be addressed in the DMP document you do provide
  \item diagrammatic workplan (one page)
  \item management structure (one page)
  \item narrative CV (up to two pages per staff member)
  \item community letters of support or demand (up to 10)  
  \item proposal cover letter.
\end{itemize}

\tableofcontents

\newpage

% \maketitle

\section{Case for support}

\ii Instructions and guidance:
  \url{https://www.ukri.org/wp-content/uploads/2021/09/BBSRC-150921-FundingOpp-BioinformaticsBiologicalResourcesFund-GuidanceDocument.pdf}

\iibf Assessment criteria.

\ii \url{https://www.ukri.org/wp-content/uploads/2021/09/BBSRC-150921-FundingOpp-BioinformaticsBiologicalResourcesFund-AssessmentCriteria.pdf}

\iibf Scientific quality and strategic relevance of the resource.

\ii Including:
\begin{itemize}[nosep]
    \item the extent to which the resource meets the highest international standards in resource provision in its field
    \item how well the resource is demonstrated to be unique or complementary to other similar existing resources
    \item the extent to which the resource addresses the research and policy priority areas of BBSRC.
\end{itemize}

\iibf Cost effectiveness, particularly considerations for long-term
sustainability beyond BBSRC funding.

\ii The extent to which:
\begin{itemize}[nosep]
    \item the resource delivers value for money relative to the anticipated scientific gains it will provide
    \item long term sustainability options have been considered, addressed, and planned where appropriate, particularly for existing resources.
\end{itemize}

\iibf Potential for economic and social impact beyond the academic
community.

\ii Including:
\begin{itemize}
    \item the extent to which the outputs from the resource will contribute to knowledge and potential for economic return or social impact
    \item how well the proposal has outlined methods of engagement and measures of success in developing milestones and timelines of associated activities.
\end{itemize}

\iibf Fit to the scope.

\ii How well the proposal addresses the scope of the opportunity.

\iibf Assessment criteria adapted to new or existing resources

\ii To allow for a more nuanced assessment between new and existing
resources, the use of the following assessment criteria will be
adapted accordingly.

\ii For new resources, these criteria will assess the ‘plans, potential, and promise’ of the resources.

\iibf Quality of the overall arrangements for resource management, advisory functions, as well as user access and engagement

\ii Including:
\begin{itemize}
    \item the extent to which the proposal has evidenced or planned interaction with relevant users and the broader research community to ensure the aims of the resource are realised and there is sufficient uptake and continued development
    \item the extent to which adequate user access arrangements have been discussed and considered
    \item the set-up of project management and advisory structures of the resource to ensure longevity in delivering the resource to a broad user base.
\end{itemize}

\iibf Need or demand, and potential benefit to the UK academic research community

\ii The extent to which:
\begin{itemize}
    \item the community has demonstrated demand for the proposed resource, relative to the total community size (in particular, proposals for new resources should have consulted their prospective community prior to application)
    \item the proposed resource will deliver and benefit the wider BBSRC community indicating how the proposed resource will help to deliver high-quality research.
\end{itemize}

\newpage
\ii Section headings taken from guide.  We can omit some if clearly not necessary, but should do so with caution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Background to the Resource}
\ii Introduction of the proposed resource, including its academic and wider 
economic and societal context.

\ii Overview of past and current resource(s) in the subject area in both the UK 
and abroad, including any alternative community resources currently 
available. You should indicate the community size of the intended resource 
and how this relates to the field in which it operates.

Experimental control and neural data analysis are currently highly
disassociated.  Most neuroscientists collect their data and only afterward
analyze it. This split has had dramatic consequences for the progress of
neuroscience. First, most experimental control methods currently do not use
advanced data analysis methods to enhance their functionality. Second, the vast
majority of neural data analysis methods are offline and do not provide key
functionality needed for online experimental control. Using advanced data
analysis methods in experimental control will improve the sophistication of
experiments that are currently possible, and generate new demands for neural
data analysis methods that will generate advances in the field.

A central goal of the software development proposed here is to address this
problem, by providing Bonsai~(an experimental control software ecosystem;
Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art online (and offline) data
analysis methods (Figure~\ref{fig:proposedBonsaiExtensions} and 
Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}).
%
It is unfortunate that neuroscientists currently do not have software that
allows them to easily compare the performance of different data analysis
methods for the characterization of their datasets. The second central goal
of this proposal is to address this limitation by adding to Bonsai a generic
mechanism for comparing the performance of different methods for the analysis
of specific datasets (Section~\ref{sec:comparisonOfMultipleMethods}).

\begin{figure}
    \begin{center}
        \includegraphics[width=7in]{figures/bonsai.pdf}
        \label{fig:bonsai}
    \end{center}
\end{figure}

\setcounter{figure}{1}

\begin{figure}
    \begin{center}
        \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
        \caption{Proposed extensions to Bonsai}
        \label{fig:proposedBonsaiExtensions}
    \end{center}
\end{figure}

Currently Bonsai has its own package manager, and its community of users have
already extended Bonsai's functionality with several contributed packages
(e.g., \ldots \textcolor{red}{Goncalo please reference the best contributed
packages developed by the Bonsai community}). However, these packages need to
be written in C\#, while most current neuroscience data analysis methods are
written in Python, R or Matlab.  We propose to add to Bonsai capabilities to
communicate with software written in these languages
(Section~\ref{sec:bonsaiPythonMatlabCommunication}).  These capabilities will
allow a large number of neuroscience data analysis methods to be easily
integrated into the Bonsai ecosystem and will provide Bonsai users a large
repertoire of advance machine learning methods for their behavioral and neural
data analysis.

Bonsai is an excellent tool for reproducible data acquisition and
experimental control. An experiment implemented in Bonsai can be replicated in
a new laboratory by just sharing a Bonsai configuration file. With the addition
of the proposed machine learning functionality Bonsai will extend this
reproducibility to the domain of data analysis
(Section~\ref{sec:reproducibility}).

\subsubsection{Bonsai}
\label{sec:bonsai}

Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
programming language that emphasizes performance, flexibility, and ease-of-use,
allowing scientists with no previous programming experience to quickly develop
their own high-performance data acquisition and experimental control systems
(Figure 1c). Bonsai combines a high-level event algebra for data streams with
an integrated development environment (IDE) and a strong library of plugins
supporting multiple hardware and software packages used by the biomedical
research community (Figures 1a, 1b).

Bonsai has a large user base in the systems neuroscience community.
Quantifications of this user base are provided in Figures 1d-1f.  In the last
year alone, we estimate more than 1,000 new users have started to incorporate
Bonsai into their experimental protocols across the world. The surprising rate
of adoption of Bonsai in non-programmer experimental labs highlights the need
for accessible programming tools that enable state-of-the-art technology but
also allow researchers to stay in control and change their experimental
paradigms.  Many open-source software tools are either inaccessible to
non-programmers, or too constrained to be of general use outside their narrow
domain of application. Bonsai has been successful because it has straddled this
gap to some extent.

The language has also helped to potentiate the growing wave of foundational
open hardware initiatives, such as the OpenEphys \citep{siegleEtAl17} and UCLA
Miniscope~\citep{caiEtAl16}, making it possible to quickly combine and
integrate these tools into new experiments~\citep{buccinoEtAl18}.
%
Bonsai has been adopted in large neuroscience undertakings like the
International Brain
Laboratory\footnote{\href{https://www.internationalbrainlab.com/}{https://www.internationalbrainlab.com/}}
and the Allen Institute for Brain
Science\footnote{\href{https://alleninstitute.org/what-we-do/brain-science/}{https://alleninstitute.org/what-we-do/brain-science/}}.
%
More recently, Bonsai has started to expand outside the domain of neuroscience
into biomedical research and biotechnology tool development, and even outside
academia into public outreach and education programs.

\subsection{Details of Resource}

\ii The case for support should outline the full details of the resource and 
associated work packages presented in the proposal.

\iitem Indication whether the project proposed to develop a new
resource or is in support of an existing one.

Existing if general Bonsai enhancements.  If specific to analysis we
may be able to argue for "new" if helpful, while still making
reference to Bonsai installed base.

\iitem Objectives for the proposal should be detailed including
individual measurable targets against which the outcome of the work
will be assessed. This should refer to the objectives set out in the
Je-S proposal form.

\iitem{Significant technical details for the development, maintenance or 
enhancement of the resource must be clearly outlined and indicated 
how this is of internationally exceptional quality. }

\iitem{If applicable, outline any proposed research efforts and how they 
  directly facilitate development of the resource. }

\iitem{For proposals looking to focus on maintaining status quo for an 
existing resource instead of suggesting further development, you 
should detail evidence of why significant upgrades are not required at 
this time and detail why the resource needs continued support to 
maintain world-leading functionality.}

\ii{Additional questions that may be considered: }

\iitem Does the facility begin or continue to support a growing field of 
bioscience, what is the anticipated growth and does the proposal 
adequately accommodate this?

\iitem How will the resource accelerate science within its field and
  beyond?

\iitem What would be the impact on the scientific community if the resource 
did not exist? How would this impact other, possibly dependent 
resources?

@@

\subsubsection{Low-dimensional representations of local-field potentials}
\label{sec:lowDimensionalRepresentationsOfLFPs}

Spikes are extracted from a higher-frequency range of extracellulary recorded
voltages. Local field potentials (LFPs) are computed from a lower-frequency
range and are another important signal to understand brain function. We propose
to use states inferred from LFPs by GLDS, GPFA, HMM and LFADS
(Section~\ref{sec:lowDimensionalRepresentationsOfSpikingData}) as
low-dimensional representations of the LFP.

\subsubsection{Neural decoding}
\label{sec:neuralDecoding}

In order to use low-dimensional representations of spiking activity and/or of
LFPs to guide the control of experiments, we need a decoder to optimally map
these low-dimensional representations to experimental controls.
%
We propose to implement in Bonsai several decoding/classification algorithms:
k-nearest neighbor, linear discriminative analysis, support vector machines,
random forests, artificial neural networks, naive Bayes and Gaussian process
classifiers.

\subsubsection{Comparing multiple data-analysis methods}
\label{sec:comparisonOfMultipleMethods}

We propose to add functionality to Bonsai to facilitate multi-method
comparisons in user-supplied datasets. Below we describe a comparison that we
will perform in Bonsai to asses the relative performance of the methods
described in the previous sections with neural recordings from behaving rats.
These recordings are currently being performed, to address scientific questions,
at the laboratory of Prof.~Akrami in the SWC.
%
Details of the experimental task appear in~\citet{akramiEtAl18}.
%
Briefly, in this task there are three nose ports (left, center, right). Rats
initiate a trial by inserting their nose in the center port for the duration of
the fixation period, until they hear a Go stimulus. During the fixation period
two stimuli $s_a$ and $s_b$ are presented. Rats should decide which stimuli is
louder. If $s_a$ is louder than $s_b$ the correct action is to poke the nose
into the right port in order to collect a liquid reward, and if $s_b$ is louder
than $s_a$ the correct choice is the left port.

@@

\subsubsection{Interfacing Bonsai with Python, R and Matlab data analysis software}
\label{sec:bonsaiPythonMatlabCommunication}

@@

\subsubsection{Reproducibility with Bonsai}
\label{sec:reproducibility}

@@

We will demonstrate each machine intelligence functionality added to Bonsai
(Section~\ref{sec:toBeAddedMachineIntelligenceFunctionality}) and the multiple
data-analysis methods comparisons
(Section~\ref{sec:comparisonOfMultipleMethods}) in Bonsai experiments. We will
then distribute Bonsai configuration files for users to reproduce these
demonstrations.

\subsubsection{Final remarks}
\label{sec:finalRemarks}

@@

\subsection{Community Demand}
\ii Evidence of community demand should be primarily driven from UK 
academic researchers working largely within BBSRC remit – see our 
Forward look for UK Bioscience for more detail on research areas covered by 
BBSRC. Demand from other users (such as academic communities outside 
of BBSRC remit or industrial users) may be appropriate to provide additional 
support, especially in highlighting the potential for economic, commercial or 
societal impact, but should not be the focus of the demand demonstration. 
Evidence provided should highlight examples of the high-quality science that 
the resource will underpin or has underpinned. Where possible and relevant, 
examples should be drawn from a wide research community to illustrate the 
broad impact of the resource to support high-quality internationally excellent 
science.

\ii The level of community demand should be benchmarked against other 
relevant resources and/or the size of the community. This will allow the fair 
assessment of resources with different user bases. The types of evidence 
that may be appropriate to provide will be different for new and existing 
resources.

\ii Evidence of wider consultation of the prospective community (e.g community 
surveys) is encouraged.

\iibf New Resources

\ii New resources should estimate the number of researchers who may engage and benefit from the resource. Evidence, where possible, would be of benefit and may include: 
\begin{itemize}
    \item  Datasets (or samples) in public or private repositories 
    \item Citations or acknowledgments 
    \item Gap analysis with existing resources 
    \item Pilot project uptake or feedback from potential users.  
\end{itemize}

\ii In particular, proposals for new resources should have consulted their prospective community prior to application

\iibf Existing Resources

\ii For existing resources this should include usage data of the
current resource. Data types may include:

\iitem Access requests from independent users/ sites

\iitem Citations or acknowledgements

\iitem Other public resources providing links to the resource

\iitem New major acquisitions captured by the resource

\ii{Additionally, existing resources need 
to evidence why this resource needs 
to be maintained/updated by the 
current grant. This could include: }

\iitem Survey data from users on what upgrades are needed

\iitem Evidence of an expanding 
user base, which requires 
additional resource

\iitem Recent developments in the 
field, which require upgrades 
to be integrated into the 
resource.

\subsection{User engagement}
\ii Discussion of user engagement provision should aim to answer the following 
questions:

\iitem Is there awareness of the resource within the user community?

\iitem How do you plan to develop the engagement strategy within the 
proposal timeframe to expand user engagement?

\iitem How have access mechanisms to ensure usability of the resource 
been considered?

\iitem How have user needs been incorporated into this proposal to ensure 
it is fit for purpose and will deliver on expectations?

\iibf New Resources

\ii Evidence should be provided as to how the resource plans to engage
stakeholders and ensure that the resource meets their needs and is
used by the community targeted by the resource.

\iibf Existing Resources

\ii Evidence should be provided as to whether the resource has
achieved the level of engagement it originally anticipated, and
consideration is given how the additional investment would change
this.

\subsection{Long term sustainability planning}
\ii{
In addition, the case for support should outline considerations for the long-
term sustainability of the resource beyond UKRI-BBSRC funding, as well as 
the true cost of running and maintaining the resource in question. 
The proposal should include: }

\iitem Cost recovery plans, where appropriate, or an explanation why
not if not viable. Evidence of clear business planning with a focus on
at least partial cost recovery is required, especially when applying
as an existing resource.

\iitem Details for alternative support plans, aside from UKRI-BBSRC 
funding.

\iitem The level of support the resource is projected to require for expected 
maintenance and/or subsequent maturation/enhancement activities. 
Clear arguments as to why UKRI-BBSRC should support the resource now 
should be provided, if other cost recovery and support plans are deemed 
unsuitable.

\subsection{Potential for economic and societal impact}
\ii{Outline how the outputs of the proposed resource will contribute to 
knowledge and how this may have the potential for economic return or 
societal benefits. Impact activities should be integrated into appropriate 
sections of the case for support, not presented as an independent work 
package. }

\iitem All proposals are expected to demonstrate clear plans with recorded 
milestones and timelines for associated activities to develop 
economic, commercial and societal impacts.  
August 2021 7

\iitem Methods of engagement and measures of success should be outlined 
including how these will be regularly reviewed throughout the project 
in order to deliver the most impact.

\iitem Any planned activities should be fully justified within the Justification 
of Resources attachment

\subsection{Track Record}
\ii{
The majority of the track record relevant to the project should be located 
within the Narrative CV and should not be repeated within the case for 
support. You may, however, want to describe: }

\iitem Track record of the team working together

\iitem Specific expertise available at the host organisation and any 
proposed partner organisations to enable the successful delivery of 
the project.

\iitem The specific role of each applicant and collaborator in the project

\bibliographystyle{apalike}
\bibliography{bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses}

\end{document}

>>>>

We intend to improve an \textbf{existing resource}, the Bonsai ecosystem. We
propose to create software infrastructure to facilitate the integration into it
of existing advanced data analysis methods
(Section~\ref{sec:softwareInfrastructure}).
%
To demonstrate the use of this infrastructure, and to provide Bonsai users with
an initial set of advanced machine intelligence neural data analysis methods,
we will add to the Bonsai ecosystem machine learning functionality to control
neuroscience experiments and to analyse data generated by them
(Section~\ref{sec:functionality}).

This initial set of methods should demonstrate to Bonsai users, and to
machine-learning methods developers, the potential of embedding
machine-learning functionality into the experimental control loop. We propose
user engagement activities to expose both of these groups of people to this
potential (Section~\ref{sec:userEngagement}).  When data-analysis method
developers recognise this potentials, they will become interested in
contributing their methods to the Bonsai ecosystem, ensuring the
\textbf{long-term sustainability} of the proposed resource.

Adding machine learning functionality to the experimental control loop will
make possible the implementation of experiments with unprecedented levels of
control. Moreover, equipping experimental scientists without programming
experience with machine learning tools will considerably increase the
application of machine learning to the biosciences, which most probably will
translate into groundbreaking scientific discoveries.

\subsection{Software infrastructure}
\label{sec:softwareInfrastructure}

\begin{comment}
Bonsai does not currently provide machine learning functionality. For instance,
the close-loop experimental control depicted in
Figure~\ref{fig:bci} cannot be currently implemented in
Bonsai. It represents the control of an experiment where a subject guides the
mouse position in a computer screen with her brain activity (Section~\ref{}).
Nodes highlighted in colour require machine intelligence functionality currently
not available in Bonsai.
%
\end{comment}

We propose to create software infrastructure to add machine intelligence
functionality to the Bonsai ecosystem in general
(Section~\ref{sec:generality}) and extensible ways
(Section~\ref{sec:extensibility}).

\subsubsection{Generality}
\label{sec:generality}

The number of current machine learning algorithms is very large and it is
continuously increasing. We intend to build a small number of machine
intelligence operators in Bonsai that can be configured to provide the
functionality of most machine learning algorithms.

For this, we will build on abstractions used in the \texttt{sickit-learn}
library. The success of this library is, to a large extent, due to its simple
and general application programming interface \citep[API;][]{buitinckEtAl13}.
%
This API consists of three complementary interfaces: an estimator interface for
building and fitting models, a predictor interface for making predictions and a
transformer interface for converting data.
%
Using only these three interfaces \texttt{scikit-learn} provides
implementations of a large number of machine learning algorithms for
dimensionality reduction, clustering, classification and regression.

Inspired by \texttt{scikit-learn}, Bonsai will have three types of operators:
estimators, predictors and transformers. Each of these operators will be
instantiated by objects implementing the corresponding interface.
%
For example, consider the close-loop system for the control of a computer
cursor with neural activity represented in Figure~\ref{fig:bci}a. The Bonsai
graphical program for this system is given in Figure~\ref{fig:bci}b. The
objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
GPFAlatentsExtractor will implement the transformer interface, and the object
instantiating the TimeSeriesForestClassifier will implement the predictor
interface (Figure~\ref{fig:bci}b, top).
%
The parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter will
be set manually by the user, as is commonly done in practice. The parameters of
the GPFAlatentsExtractor will be learnt offline in an unsupervised way in a
second pipeline (Figure~\ref{fig:bci}b, middle). Finally, the parameters of the
TimeSeriesForestClassifier will also be learnt offline, but in a supervised
way, in a third pipeline (Figure~\ref{fig:bci}b, bottom).

\begin{figure}
   \begin{center}
       \includegraphics[width=4in]{figures/bci.png}
       \caption{Close-loop system for controlling a computer cursor with brain activity}
     \label{fig:bci}
   \end{center}
\end{figure}

\begin{comment}
    The spike detector transforms membrane
potentials into spike sequences. The spike sorter transforms spikes sequences
into lists of spikes, one list per neuron. The latents extractor transforms
multi-neuron lists of spikes into latent trajectories summarising the
multi-neuron spiking activity.  The neural decoder predicts the next cursor
position from the latent trajectories.
\end{comment}

\begin{comment}
middle, whose output will set
As in this library,
there will be three type of machine intelligence operators in Bonsai:
%
estimator operators for building and fitting models, predictor operators for
making predictions and transformer operators for converting data. These
operators will be implemented by objects exposing the corresponding interfaces.

The estimator interface exposes a fit method for learning a model from training
data.  This method is called with training data (e.g., supplied as two arrays
X\_train and y\_train in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data.  All supervised and unsupervised learning algorithms (e.g., for
classification, regression or clustering) will be offered as objects
implementing this interface. Machine learning tasks like feature extraction,
feature selection or dimensionality reduction will also be provided as
estimators.

The predictor interface extends the notion of an estimator by adding a predict
method that takes an array X\_test and produces predictions for X\_test, based
on the learned parameters of the estimator (we call the input to predict
“X\_test” in order to emphasise that predict generalises to new data). In the
case of supervised learning estimators, this method typically returns the
predicted labels or values computed by the model.

Since it is common to modify or filter data before feeding it to a learning
algorithm, the transformer interface defines a transform method. It takes as
input some new data X\_test and yields as output a transformed version of
X\_test. Preprocessing, feature selection, feature extraction and
dimensionality reduction algorithms will be provided as transformers in Bonsai.

Bonsai will provide the ability to compose new
estimators from several base estimators. Composition mechanisms can be used
to combine typical machine learning workflows into a single object which is
itself an estimator, and can be employed wherever usual estimators can be used.
Composition of estimators will be done with Pipeline objects that chain multiple estimators into a single one. This is useful
since a machine learning workflow typically involves a fixed sequence of processing steps (e.g., feature extraction, dimensionality reduction, learning and making
predictions), many of which perform some kind of learning. A sequence of N such
steps can be combined into a Pipeline if the first N-1 steps are transformers;
the last can be either a predictor, a transformer or both.

\end{comment}

\subsubsection{Extensibility}
\label{sec:extensibility}

\textcolor{red}{I will start this section by saying that we are not proposing
to implement machine learning algorithms, as in \texttt{scipy.learn}, but to
integrate into Bonsai existing machine intelligence algorithms.}

Currently Bonsai has its own package manager, and its community of users have
already extended Bonsai's functionality with several contributed packages
(e.g., \ldots). However, these packages need to be written in C\#, while most
current data analysis methods are written in Python, R or Matlab. We propose to
add to Bonsai capabilities to communicate with software written in these
languages.

Fortunately, there exist software that allow C\# program
to call and be called by programs in these other languages. For the
communication of C\# programs with Python we will use
Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
with R programs we will use
R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
and with Matlab we will use the built in Matlab .NET
interface\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
In cases where these software cannot support some type of communication between
C\# and another language we will use the messaging library
ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.

Using the previous software, we will build C\# wrappers around machine
intelligence programs written in other languages, so that these programs
implement the Estimator, Predictor or Transformer interfaces and can be
integrated into Bonsai as machine intelligence operators.

\subsection{Machine intelligence functionality to be added to Bonsai}
\label{sec:functionality}

We propose to initially add to Bonsai machine intelligence functionality to
support neuroscience experiments because, currently, neuroscience is the
research area with the largest number of Bonsai users. Within neuroscience, we
propose to add functionality for video-based behavioural analysis
(Section~\ref{sec:videoBasedBehavioralAnalysis}) and for running
brain-computer-interface experiments (Section~\ref{sec:bci}), since these tasks
are relevant to many Bonsai users, and recent advances in machine learning
will greatly improve their implementation in Bonsai.

@@

\paragraph{Low-dimensional representations of local-field potentials}
% \label{sec:lowDimensionalRepresentationsOfLFPs}

Spikes are extracted from a higher-frequency range of extracellularly recorded
voltages. Local field potentials (LFPs) are computed from a lower-frequency
range and are another important signal to understand brain function. We propose
to use states inferred from LFPs by GLDS, GPFA, HMM and LFADS
(Section~\ref{sec:lowDimensionalRepresentationsOfSpikingData}) as
low-dimensional representations of the LFP.

\paragraph{Neural decoding}
% \label{sec:neuralDecoding}

In order to use low-dimensional representations of spiking activity and/or of
LFPs to guide the control of experiments, we need a decoder to optimally map
these low-dimensional representations to experimental controls.
%
We propose to implement in Bonsai several decoding/classification algorithms:
k-nearest neighbour, linear discriminative analysis, support vector machines,
random forests, artificial neural networks, naive Bayes and Gaussian process
classifiers.

\paragraph{Comparing multiple data-analysis methods}
% \label{sec:comparisonOfMultipleMethods}

We propose to add functionality to Bonsai to facilitate multi-method
comparisons in user-supplied datasets. Below we describe a comparison that we
will perform in Bonsai to asses the relative performance of the methods
described in the previous sections with neural recordings from behaving rats.
These recordings are currently being performed, to address scientific questions,
at the laboratory of Prof.~Akrami in the SWC.
%
Details of the experimental task appear in~\citet{akramiEtAl18}.
%
Briefly, in this task there are three nose ports (left, centre, right). Rats
initiate a trial by inserting their nose in the centre port for the duration of
the fixation period, until they hear a Go stimulus. During the fixation period
two stimuli $s_a$ and $s_b$ are presented. Rats should decide which stimuli is
louder. If $s_a$ is louder than $s_b$ the correct action is to poke the nose
into the right port in order to collect a liquid reward, and if $s_b$ is louder
than $s_a$ the correct choice is the left port.

@@

\begin{comment}

\subsection{Interfacing Bonsai with Python, R and Matlab data analysis software}
\label{sec:bonsaiPythonMatlabCommunication}

@@

\end{comment}

\subsection{Reproducibility with Bonsai}
\label{sec:reproducibility}

@@

We will demonstrate each machine intelligence functionality added to Bonsai
(Section~\ref{sec:functionality}) and the multiple
data-analysis methods comparisons
(Section~\ref{sec:comparisonOfMultipleMethods}) in Bonsai experiments. We will
then distribute Bonsai configuration files for users to reproduce these
demonstrations.

\subsection{Measurable targets}
\label{sec:measurableTargets}

We propose to assess the progress of the proposed project using the following
measurable targets:

\begin{comment}

    \item[implementation of software infrastructure]
        (Section~\ref{sec:softwareInfrastructure}) includes the
        implementation of the machine abstractions, described in
        Section~\ref{},
        and of the capabilities to allow Bonsai communicate with
        software written in other languages, described in Section~\ref{}. We
        will consider this target completed when we can implement a simple BCI
        application, with the architecture depicted in Figure~\ref{}, with
        machine learning functionality provided by software in Python, Matlab
        and R.

    \item[implementation of functionality] (Sections~\ref{sec:functionality})
        includes the integration into Bonsai of machine learning software for
        video-based behavioral analysis (Section~\ref{}) and for
        brain-computer-interface applications (Section~\ref{}). Completion of
        this target requires that Bonsai can perform each of these tasks
        separately, in conjunction with software written in Python, R and
        Matlab.

    \item[internal testing] multiple groups at the SWC use Bonsai to control
        their experiments.

        Bonsai Github repository and the success use by Bonsai users

    \item Bonsai users use the implemented machine learning functionality. We
        propose to evaluate this target by checking in the Bonsai user forum
        the frequency with which a novel machine learning functionality is
        discussed, and by tracking publications using this functionality.

    \item method developers integrate their methods into the Bonsai ecosystem.
        The integration of machine learning functionality into Bonsai will use
        Bonsai's package manager. We propose to use this package manager to tack
        the number of machine intelligence methods integrated into Bonsai.

\end{comment}

\subsection{Final remarks}
\label{sec:finalRemarks}

@@
